### 2016-04-29 writing/latex/coincidence.tex
Getting closer

\chapter[Coincidence Classification]{Coincidences amongst families of mesh patterns and classical patterns}\label{chap:coincs}
One interesting question to ask about permutation patterns considers when a
pattern may be avoided by, or contained in, arbitrary permutations. Two
patterns \(\pi\) and \(\sigma\) are said to be \emph{coincident} if the set
of permutations that avoid \(\pi\) is the same as the set of permutations that
avoid \(\sigma\). This extends to sets of patterns as well as single patterns.

We consider the avoidance sets, \(\av{p,q}\) where \(p\) is a classical pattern
of length \(3\) and \(q\) is a mesh pattern of length \(2\) in order to
establish some rules about when these two sets give the same avoidance set.
We fix \(p\) in order to define the equivalences, we say that \(p\) is the
\emph{Dominating Pattern}. We fix \(p\) in order to calculate the coincidences.
We never consider occurrences with different \emph{Dominating Patterns}.

We first define some operations on mesh patterns.
\begin{definition}
Given a pattern \(p\), let \(\addp[D]{p}{(a,b)}\) be the operation that returns
a mesh pattern equivalent to placing a point in the center of box \((a,b)\) in
\(p\), with shading defined by \(D\subseteq\{N,E,S,W\}\).
\end{definition}
The set \(D\) defines the shading by indicating that the boxes in the cardinal
directions in \(D\) next to the point are shaded in the resulting pattern.
Since there is no ambiguity we let \(\addpe[D]\) be equivalent to \(\addp[D]{\varepsilon}{(0,0)}\).
This operation fails if the box \((a,b)\) is in the mesh set of \(p\).

\begin{example}
    The result of adding a single point to the empty permutation for each cardinal direction.
 \begin{align*}
     \addpe[\{N\}] = \textpattern{}{1}{0/1,1/1} &\qquad\addpe[\{E\}] =\textpattern{}{1}{1/1,1/0} \qquad \\
     \addpe[\{S\}] = \textpattern{}{1}{0/0,1/0} &\qquad\addpe[\{W\}] =\textpattern{}{1}{0/0,0/1} \qquad
\end{align*}
    A more complex example for \texttt{add\_point}
\begin{equation*}
\addp[\{N,E\}]{\pattern{}{2,1,3}{0/3}}{(2,3)} = \pattern{}{2,1,4,3}{0/3,0/4,2/4,3/4,3/3}
\end{equation*}
\end{example}

\begin{definition}
    Given a pattern \(p\), define \(\addd{p}{(a,b)}\), and \\ \(\addi{p}{(a,b)}\),
    as the operations that return a mesh pattern equivalent to placing an
    decrease, or increase, in the center of box \((a,b)\) in \(p\).
\end{definition}

\begin{example}
    \begin{align*}
        \addie = \pattern{}{1,2}{0/1,1/0,1/1,1/2,2/1}\\
        \addde = \pattern{}{2,1}{0/1,1/0,1/1,1/2,2/1}
    \end{align*}
\end{example}

@@

It can be easily seen that in order to classify set equivalences one need only
consider coincidences within the family of mesh patterns with the same underlying
classical pattern, this is due to the fact that \(\perm{2,1}\in\av{\mperm{1,2}{R}}\)
and \(\perm{1,2}\in\av{\mperm{2,1}{R}}\) for all mesh-sets \(R\). So
\(\av{\mperm{1,2}{R}}\setminus \av{\mperm{2,1}{S}}\neq \emptyset\ \forall
R,S\in [0,2]\times [0,2]\), and hence the sets are disjoint.

We know that there are a total of \(512\) mesh-sets for each underlying classical
pattern. By use of the previous results of the Shading
Lemma\cite[Lemma~3.11]{DBLP:journals/combinatorics/HilmarssonJSVU15},
Simultaneous Shading Lemma\cite[Lemma~7.6]{DBLP:journals/corr/ClaessonTU14}, and
one special case, we can reduce the number of equivalence classes to \(220\).

By discussion of a number of rules we will show that the number of coincidence
classes follows the values shown in  \TableRef{tab:domclasses}.

@@

\section{Equivalence classes of Av(\{321, (21, \textit{R})\}).}
Through experimentation we discover that there are a total of 29 equivalence
classes of mesh patterns with underlying classical pattern \(\perm{2,1}\).

@@

\begin{note}
    \label{not:downcmesh}
    Let \(R^\prime \subseteq R\). Then any occurence of \((\tau, R)\) in a permutation
    is an occurence of \((\tau,R^\prime)\).
\end{note}
\begin{proof}
    Assume without meaningful loss of generality that \(R^\prime\) is a proper
    subset of \(R\).

    Consider an occurence of \((\tau, R)\) in a permutation \(\sigma\), obviously
    this corresponds to an occurrence of \(\tau\) in \(\sigma\). Now consider
    the mesh sets \(R\) and \(R^\prime\), since \(R^\prime \subseteq R\) then
    there are more restrictions on where points are in an occurrence of \((\tau,R)\).
    Namely, for every shaded box in \(R\) the corresponding region in \(\sigma\)
    must contain no points, since \(R^\prime\) has less shading than \(R\) there
    exists a region in the occurence of \((\tau,R)\) in \(\sigma\) that is
    now devoid of restrictions. However, by removing restrictions we cannot
    make an occurence become not an occurence, and therefore the same occurrence
    of \(\tau \) in \(\sigma\) is now an occurence of \((\tau, R^\prime)\).
\end{proof}
\begin{proof}[Proof of \PropositionRef{prop:dom1}]
    We need to prove that \(\av{\{\pi,m_1\}} = \av{\{\pi,m_2\}}\).

@@

By using this rule we completely capture the equivalence classes of
\(\av{\{\perm{3,2,1}, \mperm{2,1}{R}\}}\).

@@

        \caption{Visual Depiction of first dominating pattern rule.}
        \label{fig:rule1}
    \end{center}
\end{figure}

\section{Equivalence classes of Av(\{231, (21, \textit{R})\}).}
By application of \PropositionRef{prop:dom1} we obtain \(43\) equivalence
classes. Experimentation shows that there are in fact \(39\) equivalence
classes, for example the following two patterns are coincident in
\(\av{\perm{2,3,1}}\) but this is not explained by \PropositionRef{prop:dom1}.
\begin{equation*}
    \pattern{}{2,1}{0/0,0/2,1/0,2/1} \text{ and } \pattern{}{2,1}{0/0,0/2,1/0,1/1,2/1}
\end{equation*}

Consider the box \((1,1)\), this is the box that we need to find reasoning
to allow shading. If this box were to be empty then we could shade it freely,
consider what happens if there are points in this box.  If there is any
increase in this box then we would have an occurrence of \(\perm{2,3,1}\),
however, since we are in \(\av{\perm{2,3,1}}\) this is not possible. This box
must contain a decreasing subsequence. This gives rise to the following lemma:

@@

Going back to our example mesh patterns
\begin{equation*}
    \begin{tikzpicture}[scale=\picscale]
        \modpattern{}{2,1}{0/0,0/2,1/0,2/1}
        \filldraw (1.66,1.33) circle (3pt);
        \filldraw (1.33,1.66) circle (3pt);
    \end{tikzpicture}
\end{equation*}
We know that the box \((1,1)\) contains an decreasing subsequence. The top point
of this decrease can be chosen to act as the second point in the mesh pattern,
and therefore there are no points between the first point and the new second point.
Hence, we can shade this box as it is guaranteed to be empty. This shows that our
two example patterns are coincident.

@@

By taking the First Dominating Pattern Rule and the Second Dominating Pattern
rule together coincidences of classes of the form \(\av{\{\perm{1,3,2},\mperm{2,1}{R}\}}\)
are fully explained, obtaining 39 equivalence classes of mesh patterns.

@@

For example the patterns
\begin{equation*}
    \pattern{}{1,2}{0/2,2/0,2/1} \text{ and } \pattern{}{1,2}{0/2,1/0,2/0,2/1}
\end{equation*}
are experimentally coincident. This coincidence is not explained by our rules,
it is necessary to attempt to capture these coincidences by establishing more
rules.

In order to rigorously establish this coincidence we need to consider what would
happen if we were to choose a point in the box \((1,0)\) that we would like to shade.
In order to get a chance of getting the shading we want consider choosing the
rightmost point in the box.
This gives us the following mesh pattern.
\begin{equation*}
    \pattern{}{2,1,3}{0/3,2/0,2/1,3/0,3/1,3/2}
\end{equation*}
By application of the \PropositionRef{prop:dom1} we then achieve the following
mesh pattern
\begin{equation*}
    \pattern{2,3}{2,1,3}{0/3,1/3,2/0,2/1,3/0,3/1,3/2}
\end{equation*}
If we look at the highlighted points we see that the subpattern is an occurrence
of the mesh pattern that we originally desired. This gives rise to the following
rule:

\begin{proposition}[Third Dominating Pattern Rule]
    \label{prop:dom3}
    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
    and a dominating classical pattern \(\pi = (\pi,\emptyset)\)\footnote{The
    permutation \(\pi\) may be the empty permutation}, the sets
    \(\av{\{\pi,m_1\}}\) and \(\av{\{\pi,m_2\}}\) are coincident if
    \begin{enumerate}
        \item \(R_1 \triangle R_2 = \{(a,b)\}\)
        \item\label{prop:dom3:condocc} \(\addp[D]{\mperm{\sigma}{R_1}}{(a,b)}\)
            is coincident with a mesh pattern containing an occurrence of
            \(\mperm{\sigma}{R_2}\).
    \end{enumerate}
\end{proposition}
\begin{proof}
    We need to prove that \(\av{\{\pi,m_1\}} = \av{\{\pi,m_2\}}\).

@@

The patterns
\begin{equation*}
    \pattern{}{1,2}{0/0, 0/2, 1/1, 1/2, 2/1} \text { and } \pattern{}{1,2}{0/0, 0/1, 0/2, 1/1, 1/2, 2/1}
\end{equation*}
are equivalent in \(\av{\perm{3,2,1}}\). (There are 3 symmetries of these
patterns that are also equivalent to each other by the same reasoning.)

Consider the contents of the box \((0,1)\), by \LemmaRef{lem:incdec} it
must contain an increasing subsequence. If there is only one point in the box
we can choose this to replace the \(1\) in the mesh pattern to get the
required shading. If there is more than one point then choosing the
two leftmost points give us the following mesh pattern.
\begin{equation*}
    \pattern{3,4}{2,3,1,4}{0/0, 0/1, 0/2, 0/3, 0/4,
                            1/0, 1/1, 1/2, 1/3, 1/4,
                            2/0, 2/1,2/2, 2/4,
                            3/1, 3/2, 3/3, 3/4,
                            4/1, 4/2, 4/3}
\end{equation*}
Where the two highlighted points are the original two points.
Now if we take the other two points as the points in our
mesh permutation then we get an occurrence of the pattern
we originally desired, and hence the two patterns are coincident.

>>>>

\chapter[Coincidence Classification]{Coincidences amongst families of mesh patterns and classical patterns}\label{chap:coincs}
One interesting question to ask about permutation patterns considers when a
pattern may be avoided by, or contained in, arbitrary permutations. Two
patterns \(\pi\) and \(\sigma\) are said to be \emph{coincident} if the set
of permutations that avoid \(\pi\) is the same as the set of permutations that
avoid \(\sigma\) \ie \(\av{\pi} = \av{\sigma}\). This extends to sets of patterns as well as single patterns.

We consider the avoidance sets, \(\av{p,q}\) where \(p\) is a classical pattern
of length \(3\) and \(q\) is a mesh pattern of length \(2\) in order to
establish some rules about when these two sets give the same avoidance set.
We fix \(p\) in order to define the equivalences and say that \(p\) is the
\emph{dominating pattern}.

We first define some operations on mesh patterns.
\begin{definition}
Given a mesh pattern \(p\), let \(\addp[D]{p}{(a,b)}\) be the operation that returns
a mesh pattern equivalent to placing a point in the center of box \((a,b)\), where \((a,b)\) is not the
mesh set of \(p\) in
the figure of \(p\), with shading defined by \(D\subseteq\{N,E,S,W\}\).
The set \(D\) defines the shading by indicating that the boxes in the cardinal
directions in \(D\) next to the point are shaded in the resulting pattern.
\end{definition}
Since there is no ambiguity we let \(\addpe[D]\) be equivalent to \(\addp[D]{\varepsilon}{(0,0)}\).

\begin{example}
    The result of adding a single point to the empty permutation for each cardinal direction.
 \begin{align*}
     \addpe[\{N\}] = \textpattern{}{1}{0/1,1/1} &\qquad\addpe[\{E\}] =\textpattern{}{1}{1/1,1/0} \qquad \\
     \addpe[\{S\}] = \textpattern{}{1}{0/0,1/0} &\qquad\addpe[\{W\}] =\textpattern{}{1}{0/0,0/1} \qquad
\end{align*}
    A more complex example for \texttt{add\_point}
\begin{equation*}
\addp[\{N,E\}]{\pattern{}{2,1,3}{0/3}}{(2,3)} = \pattern{}{2,1,4,3}{0/3,0/4,2/4,3/4,3/3}
\end{equation*}
\end{example}
Formally, given a mesh pattern \(p = (\pi,R)\) \(\addp[D]{p}{(a,b)}\) gives a mesh
pattern \(p^\prime = (\pi^\prime,R^\prime)\) defined by
\begin{equation*}
    \pi^\prime(i) = \begin{cases}
    \pi(i) & \text{if } i\neq a+1 \text{ and } \pi(i)<b\\
    \pi(i)+1 & \text{if } i\neq a+1 \text{ and } \pi(i)>b\\
    b+1 & \text{if } i = a+1
\end{cases}
\end{equation*}
and
\begin{equation*}
        R^\prime = \bigcup_{(i,j)\in R}r((i,j))
\end{equation*}
Where \(r((i,j))\) is defined by
\begin{equation*}
    r((i,j)) =
    \begin{cases}
        \{(i,j)\} & \text{if } i<a, j<b\\
        \{(i,j),(i,j+1)\} & \text{if } i<a, j=b\\
        \{(i,j+1)\} & \text{if } i<a, j>b\\
        \{(i,j),(i+1,j)\} & \text{if } i=a, j<b\\
        \{(i,j+1),(i+1,j+1)\} & \text{if } i=a, j>b\\
        \{(i+1,j)\} & \text{if } i>a, j<b\\
        \{(i+1,j),(i+1,j+1)\} & \text{if } i>a, j=b\\
        \{(i+1,j+1)\} &\text{if } i>a, j>b\\
    \end{cases}
\end{equation*}

If the shading set \(D\) is non-empty we can modify definition of the directions
slightly
\begin{equation*}
    \begin{aligned}
        N &= \{(a,b+1),(a+1,b+1)\} \\
        E &= \{(a+1,b),(a+1,b+1)\} \\
        S &= \{(a,b),(a+1,b)\} \\
        W &= \{(a,b),(a,b+1)\}
    \end{aligned}
\end{equation*}
And we add the union of the sets in \(D\) into the mesh set \(R^\prime\).

\begin{definition}
    Given a pattern \(p\), define \(\addd{p}{(a,b)}\), and \\ \(\addi{p}{(a,b)}\),
    as the operations that return a mesh pattern equivalent to placing an
    decrease, or increase, in the center of box \((a,b)\), where \((a,b)\) is not
    in the mesh set of \(p\), in \(p\).
\end{definition}

\begin{example}
    \begin{align*}
        \addie = \pattern{}{1,2}{0/1,1/0,1/1,1/2,2/1}\\
        \addde = \pattern{}{2,1}{0/1,1/0,1/1,1/2,2/1}
    \end{align*}
\end{example}
We can define these formally as with \(\addp[D]{p}{(a,b)}\). Considering
only adding the ascent, adding a descent is very similar
\begin{equation*}
    \addi{p}{(a,b)} = (\pi^\prime, R^\prime)
\end{equation*}
\begin{equation*}
    \pi^\prime(i) = \begin{cases}
    \pi(i) & \text{if } i\neq a+1,a+2 \text{ and } \pi(i)<b\\
    \pi(i)+2 & \text{if } i\neq a+1,a+2 \text{ and } \pi(i)>b\\
    b+1, b+2 & \text{if } i = a+1, a+2
\end{cases}
\end{equation*}
and
\begin{equation*}
        R^\prime = \bigcup_{(i,j)\in R}r((i,j)) \cup 
\end{equation*}
Where \(r((i,j))\) is defined by
\begin{equation*}
    r((i,j)) =
    \begin{cases}
        \{(i,j)\} & \text{if } i<a, j<b\\
        \{(i,j),(i,j+1),(i,j+2)\} & \text{if } i<a, j=b\\
        \{(i,j+2)\} & \text{if } i<a, j>b\\
        \{(i,j),(i+1,j),(i+2,j)\} & \text{if } i=a, j<b\\
        \{(i,j+2),(i+1,j+2),(i+2,j+2)\} & \text{if } i=a, j>b\\
        \{(i+2,j)\} & \text{if } i>a, j<b\\
        \{(i+2,j),(i+2,j+1),(i+2,j+2)\} & \text{if } i>a, j=b\\
        \{(i+2,j+2)\} &\text{if } i>a, j>b\\
    \end{cases}
\end{equation*}

@@

It can be easily seen that in order to classify set equivalences one need only
consider coincidences within the family of mesh patterns with the same underlying
classical pattern, this is due to the fact that \(\perm{2,1}\in\av{\mperm{1,2}{R}}\)
and \(\perm{1,2}\in\av{\mperm{2,1}{R}}\) for all mesh-sets \(R\).

We know that there are a total of \(512\) mesh-sets for each underlying classical
pattern. By use of the previous results of \textcite{DBLP:journals/corr/ClaessonTU14}\footnote{
The authors use the Simultaneous Shading Lemma, a closure result and one worked out special case.
} the number of coincidence classes can be reduced to \(220\).
% the Shading
% Lemma\cite[Lemma~3.11]{DBLP:journals/combinatorics/HilmarssonJSVU15},
% Simultaneous Shading Lemma\textcite[Lemma~7.6]{DBLP:journals/corr/ClaessonTU14}, and
% one special case, we can reduce the number of equivalence classes to \(220\).

By discussion of a number of rules we will show that the number of coincidence
classes follows the values shown in  \TableRef{tab:domclasses}. The experimental
data is calculated on permutations up to length 11.

@@

\section{Coincidence classes of Av(\{321, (21, \textit{R})\}).}
Through experimentation we discover that there are at most 29 coincidence
classes of mesh patterns with underlying classical pattern \(\perm{2,1}\).

@@

\begin{note}
    \label{not:downcmesh}
    Let \(R^\prime \subseteq R\). Then any occurence of \((\tau, R)\) in a permutation
    is an occurence of \((\tau,R^\prime)\).
\end{note}
% \begin{proof}
%     Assume without meaningful loss of generality that \(R^\prime\) is a proper
%     subset of \(R\).
%
%     Consider an occurence of \((\tau, R)\) in a permutation \(\sigma\), obviously
%     this corresponds to an occurrence of \(\tau\) in \(\sigma\). Now consider
%     the mesh sets \(R\) and \(R^\prime\), since \(R^\prime \subseteq R\) then
%     there are more restrictions on where points are in an occurrence of \((\tau,R)\).
%     Namely, for every shaded box in \(R\) the corresponding region in \(\sigma\)
%     must contain no points, since \(R^\prime\) has less shading than \(R\) there
%     exists a region in the occurence of \((\tau,R)\) in \(\sigma\) that is
%     now devoid of restrictions. However, by removing restrictions we cannot
%     make an occurence become not an occurence, and therefore the same occurrence
%     of \(\tau \) in \(\sigma\) is now an occurence of \((\tau, R^\prime)\).
% \end{proof}
\begin{proof}[Proof of \PropositionRef{prop:dom1}]
    We need to prove that \(\av{\{\pi,m_1\}} = \av{\{\pi,m_2\}}\).

    Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
    % Consider a permutation \(\omega\) that contains an occurrence of \(m_2\)
    % by \NoteRef{not:downcmesh} any of these occurrences is also an occurence
    % of \(m_1\). This proves that every occurence of \(m_2\) is also an
    % occurence of \(m_1\) and therefore \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\).
    Since \(R_1\) is a subset of \(R_2\), \NoteRef{not:downcmesh} states that
    \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\)

@@

All coincidence classes of \(\av{\{\perm{3,2,1}, \mperm{2,1}{R}\}}\) can be
explained by repeated application of \PropositionRef{prop:dom1}.

@@

        \caption{Visual depiction of first dominating pattern rule.}
        \label{fig:rule1}
    \end{center}
\end{figure}

\section{Equivalence classes of Av(\{231, (21, \textit{R})\}).}
By application of \PropositionRef{prop:dom1} we obtain \(43\) equivalence
classes. Experimentation shows that there are in fact \(39\) equivalence
classes, for example the following two patterns are coincident in
\(\av{\perm{2,3,1}}\) but this is not explained by \PropositionRef{prop:dom1}.
\begin{equation*}
    m_1 = \pattern{}{2,1}{0/0,0/2,1/0,2/1} \text{ and } m_2 = \pattern{}{2,1}{0/0,0/2,1/0,1/1,2/1}
\end{equation*}

Consider an occurrence of \(m_1\) in a permutation in \(\av{\perm{2,3,1}}\), then if the region corresponding
to the box \((1,1)\) is empty we have an occurrence of \(m_2\). Otherwise, if there is any
increase in this box then we would have an occurrence of \(\perm{2,3,1}\),
however, since we are in \(\av{\perm{2,3,1}}\) this is not possible. This box
must therefore contain a decreasing subsequence. This gives rise to the following lemma:

@@

Going back to our example mesh patterns
\begin{equation*}
    \begin{tikzpicture}[scale=\picscale]
        \modpattern{}{2,1}{0/0,0/2,1/0,2/1}
        \filldraw (1.66,1.33) circle (3pt);
        \filldraw (1.33,1.66) circle (3pt);
    \end{tikzpicture}
\end{equation*}
We know that the region corresponding to the box \((1,1)\) contains a decreasing subsequence. The top point
of this decreasing subsequence can be chosen to act as the second point in the mesh pattern,
and therefore there are no points between the first point and the new second point.
Hence, we can shade this box as it is guaranteed to be empty. This shows that our
two example patterns are coincident.

@@

By taking the First Dominating Pattern Rule and the Second Dominating Pattern
Rule together coincidences of classes of the form \(\av{\{\perm{2,3,1},\mperm{2,1}{R}\}}\)
are fully explained, obtaining 39 equivalence classes of mesh patterns.

@@

For example the patterns
\begin{equation*}
    m_1 = \pattern{}{1,2}{0/2,2/0,2/1} \text{ and } m_2 = \pattern{}{1,2}{0/2,1/0,2/0,2/1}
\end{equation*}
are experimentally coincident. This coincidence is not explained by our rules,
it is necessary to attempt to capture these coincidences by establishing more
rules.

Consider an occurrence of \(m_1\) in a permutation, if the region corresponding
to the box \((1,0)\) is empty then we have an occurrence of \(m_2\). Now look at
the case when this region is not empty. Consider choosing the rightmost point in
region.
% In order to rigorously establish this coincidence we need to consider what would
% happen if we were to choose a point in the box \((1,0)\) that we would like to shade.
% In order to get a chance of getting the shading we want consider choosing the
% rightmost point in the box.
This gives us the following mesh pattern.
\begin{equation*}
    \pattern{}{2,1,3}{0/3,2/0,2/1,3/0,3/1,3/2}
\end{equation*}
By application of the \PropositionRef{prop:dom1} we then achieve the following
mesh pattern
\begin{equation*}
    \pattern{2,3}{2,1,3}{0/3,1/3,2/0,2/1,3/0,3/1,3/2}
\end{equation*}
If we look at the highlighted points we see that the subpattern is an occurrence
of the mesh pattern that we originally desired. This gives rise to the following
rule:

\begin{proposition}[Third Dominating Pattern Rule]
    \label{prop:dom3}
    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
    and a dominating classical pattern \(\pi = (\pi,\emptyset)\), the sets
    \(\av{\{\pi,m_1\}}\) and \(\av{\{\pi,m_2\}}\) are coincident if
    \begin{enumerate}
        \item \(R_1 \triangle R_2 = \{(a,b)\}\)
        \item\label{prop:dom3:condocc} \(\addp[D]{\mperm{\sigma}{R_1}}{(a,b)}\)
            is coincident with a mesh pattern containing an occurrence of
            \(\mperm{\sigma}{R_2}\).
    \end{enumerate}
\end{proposition}
\begin{proof}
    We need to prove that \(\av{\{\pi,m_1\}} = \av{\{\pi,m_2\}}\).

@@

The patterns
\begin{equation*}
    m_1 = \pattern{}{1,2}{0/0, 0/2, 1/1, 1/2, 2/1} \text { and } m_2 \pattern{}{1,2}{0/0, 0/1, 0/2, 1/1, 1/2, 2/1}
\end{equation*}
are equivalent in \(\av{\perm{3,2,1}}\). (There are 3 symmetries of these
patterns that are also equivalent to each other by the same reasoning.)

Consider the region corresponding to the box \((0,1)\) in any occurrence of \(m_1\),
in a permutation by \LemmaRef{lem:incdec} it must contain an increasing subsequence.
If there is only one point in the region we can choose this to replace the \(1\) in
the mesh pattern to get the required shading. If there is more than one point then
choosing the two leftmost points give us the following mesh pattern.
\begin{equation*}
    \pattern{3,4}{2,3,1,4}{0/0, 0/1, 0/2, 0/3, 0/4,
                            1/0, 1/1, 1/2, 1/3, 1/4,
                            2/0, 2/1,2/2, 2/4,
                            3/1, 3/2, 3/3, 3/4,
                            4/1, 4/2, 4/3}
\end{equation*}
Where the two highlighted points are the original two points.
Now if we take the other two points as the points in our
mesh permutation then we get an occurrence of the pattern
we originally desired, and hence the two patterns are coincident.

### 2016-04-29 writing/latex/introduction.tex
Getting closer

We can display a permutation on a \emph{figure} in order to give a graphical
representation of the permutation. In such a figure we let the \(x\)-axis
denote the index in the permutation, and the \(y\)-axis denotes the values of
\(\pi(x)\). The figure of the permutation \(\pi = \perm{2,3,1}\) is shown below
\begin{equation*}
    \pattern{}{2,3,1}{}
\end{equation*}
It is convenient to call the elements of the permutation \emph{points} when
referring to these figures.

@@

\begin{definition}{Order isomorphism.}
    Two sequences \(\range{\alpha_1,\alpha_2}{\alpha_n}\) and
    \(\range{\beta_1,\beta_2}{\beta_n}\) are said to be \emph{order isomorphic}
    if they share the same relative order,i.e.,\(\alpha_r<\alpha_s\) if and
    only if \(\beta_r<\beta_s\).
\end{definition}

A permutation \(\pi\) us said to \emph{contain} the permutation \(\sigma\) of
length \(k\) as a pattern (denoted \(\sigma \preceq \pi\)) if there is some
an increasing subsequence \(\range{i_1,i_2}{i_n}\) such that the sequence
\(\range{\pi(i_1),\pi(i_2)}{\pi(i_k)}\) is order isomorphic to
\(\range{\sigma(1),\sigma(2)}{\sigma(k)}\). If \(\pi\) does not contain
\(\sigma\), we say that \(\pi\) \emph{avoids} \(\sigma\).

For example the permutation \(\pi = \perm{2,4,1,5,3}\) contains the pattern
\(\sigma = \perm{2,3,1}\), since the second, fourth and fifth entries
(\(4,5,\) and \(3\)) share the same relative order as the entries of \(\sigma\).
This can be seen graphically below, the points order isomorphic to \(\sigma\)
are highlighted.
\begin{equation*}
    \pattern{2,4,5}{2,4,1,5,3}{}
\end{equation*}

@@

\section{Mesh Patterns}
Mesh Patterns were introduced by \textcite{journals/combinatorics/BrandenC11} to capture explicit expansions
for certain permutation statistics. They are a natural extension of Classical
permutation patterns. A \emph{mesh-pattern} is a pair
\begin{equation*}
    p = (\tau,R)\text{ with } \tau \in \mathfrak{S}_k \text{ and } R \subseteq
    [0,k]\times [0,k].
\end{equation*}
By this definition the empty permutation \(\varepsilon\) as a mesh pattern consisting solely of the
box \((0,0)\).

The figure for a mesh pattern looks similar to that for a classical pattern with the
addition that we shade the unit square with bottom corner \((i,j)\) for each \((i,j) \in R\):
\begin{equation*}
    \pattern{}{ 2, 1, 3 }{ 0/1, 0/2, 1/0, 1/1, 2/1, 2/2 }
\end{equation*}

We define containment, and avoidance, of the pattern \(p\) in the permutation
\(\pi\) on mesh patterns analogously to classical containment, and avoidance,
of \(\tau\) in \(\pi\) with the additional restrictions on the relative
position of the occurence of \(\tau\) in \(\pi\). These restrictions say that
the shaded regions of the figure above contain no points from \(\pi\).

@@

\begin{note}
    \label{not:classmesh}
    Classical patterns are just a subclass of mesh patterns where the mesh
    set \(R\) is empty. The classical pattern \(\pi\) can be represented
    by a mesh pattern as \((\pi,\emptyset)\).
\end{note}

In the past people have studied different classes
of permutations that can be described by mesh patterns. \textcite{babstein2000}
considered \emph{vincular} patterns (also known as \emph{generalised} or \emph{dashed} patterns),
those where two adjacent entries in the pattern must be adjacent in the permutation,
\ie entire columns are shaded in the mesh.
\textcite{MR2652101} look at classes of pattern where both columns and rows
can be shaded, these are called \emph{bivincular} patterns.
\emph{Bruhat-restricted} patterns were studied by \textcite{MR2264071} in order
to establish necessary conditions for a Schubert variety to be Gorenstein.
Mesh patterns also encompass a subset of \emph{barred} patterns introduced by
\textcite{MR2716312}.

>>>>

We can display a permutation on a \emph{figure} in order to give a graphical
representation of the permutation. In such a figure we plot the points
\((i,\pi(i))\) in a Cartesian coordinate system. The figure of the permutation
\(\pi = \perm{2,3,1}\) is shown below
\begin{equation*}
    \pattern{}{2,3,1}{}
\end{equation*}
It is convenient to call the elements of the permutation \emph{points} when
referring to these figures.

@@

\begin{definition}{Order isomorphism.}
    Two substrings \(\alpha_1\alpha_2\dotsm\alpha_n\) and
    \(\beta_1\beta_2\dotsm\beta_n\) are said to be \emph{order isomorphic}
    if they share the same relative order, \ie, \(\alpha_r<\alpha_s\) if and
    only if \(\beta_r<\beta_s\).
\end{definition}

A permutation \(\pi\) is said to \emph{contain} the permutation \(\sigma\) of
length \(k\) as a pattern (denoted \(\sigma \preceq \pi\)) if there is some
subsequence \(i_1i_2\dotsm{}i_n\) such that the sequence
\(\pi(i_1)\pi(i_2)\dotsm\pi(i_k)\) is order isomorphic to
\(\sigma(1)\sigma(2)\dotsm\sigma(k)\). If \(\pi\) does not contain
\(\sigma\), we say that \(\pi\) \emph{avoids} \(\sigma\).

For example the permutation \(\pi = \perm{2,4,1,5,3}\) contains the pattern
\(\sigma = \perm{2,3,1}\), since the second, fourth and fifth elements
(\(453\)) are order isomorphic to \(\perm{231}\).
This can be seen graphically below, the points order isomorphic to \(\sigma\)
are highlighted.
\begin{equation*}
    \pattern{2,4,5}{2,4,1,5,3}{}
\end{equation*}

@@

\section{Mesh Patterns}
Mesh Patterns were introduced by \textcite{journals/combinatorics/BrandenC11} to capture explicit expansions
for certain permutation statistics. They are a natural extension of Classical
permutation patterns. A \emph{mesh-pattern} is a pair
\begin{equation*}
    p = (\tau,R)\text{ with } \tau \in \mathfrak{S}_k \text{ and } R \subseteq
    [0,k]\times [0,k].
\end{equation*}

The set \(R\) is called the \emph{mesh-set} of the mesh pattern \(p\).
The figure for a mesh pattern looks similar to that for a classical pattern with the
addition that we shade the unit square with bottom left corner \((i,j)\) for each \((i,j) \in R\):
\begin{equation*}
    \pattern{}{ 2, 1, 3 }{ 0/1, 0/2, 1/0, 1/1, 2/1, 2/2 }
\end{equation*}

We define containment, and avoidance, of the pattern \(p\) in the permutation
\(\pi\) on mesh patterns analogously to classical containment, and avoidance,
of \(\tau\) in \(\pi\) with the additional restrictions on the relative
position of the occurence of \(\tau\) in \(\pi\). These restrictions say that
no elements of \(\pi\) are allowed in the shaded regions of the figure.

@@

Given a mesh pattern \(p=(\sigma,R)\) we say that \(\sigma\) is the \emph{underlying
classical pattern} of \(p\).
\begin{note}
    \label{not:classmesh}
    Classical patterns are a subclass of mesh patterns the classical pattern \(\pi\)
    can be represented by a mesh pattern as \((\pi,\emptyset)\).
\end{note}

In the past people have studied different classes
of permutations that can be described by mesh patterns. \textcite{babstein2000}
considered \emph{vincular} patterns (also known as \emph{generalised} or \emph{dashed} patterns),
those where two adjacent entries in the pattern must be adjacent in the permutation,
\ie \(R\) is a union of vertical strips.
\textcite{MR2652101} look at classes of pattern where both columns and rows
can be shaded, these are called \emph{bivincular} patterns.
\emph{Bruhat-restricted} patterns were studied by \textcite{MR2264071} in order
to establish necessary conditions for a Schubert variety to be Gorenstein.
Mesh patterns also encompass a subset of \emph{barred} patterns introduced by
\textcite{MR2716312}, those with only one barred letter.
